---
title: "School Accreditation Work - Post Jamovi"
author: "Elise Fashimpaur"
date: "2024-10-22"
output: html_document
---

## Packages Set-Up 
Need to include all the merged CSV's. 
```{r setup, include=FALSE}
#install.packages("dplyr")
#install.packages("readr")
#install.packages("tidyverse")
#install.packages("ggplot2")

library(dplyr)
library(readr)
library(tidyverse)
library(ggplot2)
```

## CSV's to read in

```{r}
filtered_path <- "/Users/Owner/Documents/filtered_schools_Nov_Edition.csv"
jamovi_filtered <- read_csv(filtered_path)
```

```{r}
regional_comparators <- c("Regional_Peer", "Regional_Competitor", "Willamette")
national_comparators <- c("National_Peer", "Willamette")
all_comparators <- c("Regional_Peer", "Regional_Competitor", "National_Peer", "Willamette")

comparator_colors <- c("chocolate4","grey20","grey","red")
```

## Visualizations and Tables

### Creating a Table

*Unfinished...*

```{r}
table_for_blank <- data.table(column=c("type1", "type2"))
```

### PCTPELL and C100_4

```{r}
pell_completion_all <- jamovi_filtered %>%
  ggplot(aes(x=PCTPELL_DCS*100,y=C100_4))+
  geom_jitter(alpha=0.8)+
  labs(title=element_text("4-Year Completion Rate Based on Pell Composition"),x="Percentage of Students Receiving Pell", y="Completion Rate in 4 Years")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()
pell_completion_all

pell_completion_all_linear <- jamovi_filtered %>%
  ggplot(aes(x=PCTPELL_DCS*100,y=C100_4))+
  geom_jitter(alpha=0.8)+
  labs(title=element_text("4-Year Completion Rate Based on Pell Composition"),x="Percentage of Students Receiving Pell", y="Completion Rate in 4 Years")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(method="lm",se=FALSE)+
  theme_light()
pell_completion_all_linear

pell_completion_grouped <- subset(jamovi_filtered, Comparator_Type %in% all_comparators) %>%
  ggplot(aes(x=PCTPELL_DCS*100,y=C100_4,color=Comparator_Type))+
  geom_jitter(alpha=0.8)+
  labs(title=element_text("4-Year Completion Rate Based on Pell Composition (2022-2023)"),x="Percentage of Students Receiving Pell", y="Completion Rate in 4 Years")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(method="lm",se=FALSE)+
  theme_light()
pell_completion_grouped
```

### C100_4 Over Time

```{r}
completion_regional <- ggplot(subset(jamovi_filtered, Comparator_Type %in% regional_comparators), aes(x=year_num, y=C100_4, color=INSTNM))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("4-Year Completion Rate by Year for Regional Comparators"),x="Year (2000 to 2023)", y="Completion Rate in 4 Years")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(method="lm",se=FALSE)+
  xlim(2000,2023)+
  theme_light()
  
completion_regional

completion_national <- ggplot(subset(jamovi_filtered, Comparator_Type %in% national_comparators), aes(x=year_num, y=C100_4, color=INSTNM))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("4-Year Completion Rate by Year for National Comparators"),x="Year (2000 to 2023)", y="Completion Rate in 4 Years")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(method="lm",se=FALSE)+
  xlim(2000,2023)+
  theme_light()
  
completion_national

completion_all_comparison <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=C100_4, color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("4-Year Completion Rate by Year"),x="Year (2003 to 2023)", y="Completion Rate (in 4 Years)")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))+
  xlim(2003,2023)

completion_all_comparison
```

### Grad_Debt_Mdn Over Time

```{r}
grad_debt_mdn_plot <- ggplot(data = subset(jamovi_filtered, INSTNM %in% "Willamette University"), aes(x=year_num,y=GRAD_DEBT_MDN))+
  geom_point()+
  labs(title=element_text("Graduate Debt by Year for Willamette University"),x="Year (1996 to 2023)", y="Graduate Debt")+
  theme(plot.title=element_text(hjust=0.5))+
  theme_light()
grad_debt_mdn_plot

grad_debt_comparison_tuition <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=GRAD_DEBT_MDN, color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Graduate Debt by Year"),x="Year (1996 to 2023)", y="Graduate Debt (in US Dollars, Not Accounting for Inflation)")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))+
  geom_line(aes(y=COSTT4_A,color=Comparator_Type)) #Annual Tuition Cost

grad_debt_comparison_tuition

grad_debt_comparison_full <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=GRAD_DEBT_MDN, color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Graduate Debt by Year"),x="Year (1996 to 2023)", y="Graduate Debt (in US Dollars, Not Accounting for Inflation)")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))

grad_debt_comparison_full

grad_debt_comparison_2005 <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=GRAD_DEBT_MDN, color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Graduate Debt by Year"),x="Year (2005 to 2023)", y="Graduate Debt (in US Dollars, Not Accounting for Inflation)")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))+
  xlim(2005,2023)

grad_debt_comparison_2005
```

### Diversity Percentages Over Time

```{r}
women_comparison <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=(UGDS_WOMEN*100), color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Female Enrollment Across Time"),x="Year (1996 to 2023)", y="Percentage of Women in Enrollment Composition")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))

women_comparison
```

```{r}
black_comparison <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=(UGDS_BLACK*100), color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Black Student Enrollment Across Time"),x="Year (2008 to 2023)", y="Percentage of Blacks in Enrollment Composition")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))+
  xlim(2008,2023)

black_comparison
```

```{r}
hispanic_comparison <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=(UGDS_HISP*100), color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Hispanic Student Enrollment Across Time"),x="Year (2008 to 2023)", y="Percentage of Hispanics in Enrollment Composition")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))+
  xlim(2008,2023)

hispanic_comparison
```


```{r}
asian_comparison <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=(UGDS_ASIAN*100), color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Asian Student Enrollment Across Time"),x="Year (2008 to 2023)", y="Percentage of Asians in Enrollment Composition")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))+
  xlim(2008,2023)

asian_comparison
```


```{r}
pell_comparison <- jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=(FTFTPCTPELL*100), color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("First-Year Students Receiving Pell Across Time"),x="Year (2008 to 2023)", y="Percentage of Pell Recipients in Enrollment Composition")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  scale_linetype_manual(values=c("dashed","dashed","dotdash","solid"))+
  xlim(2008,2023)

pell_comparison
```

#### Facet Wrap

```{r}
grad_debt_mdn_scatter <- ggplot(jamovi_filtered, aes(x=GRAD_DEBT_MDN, y=COMP_ORIG_YR4_RT, color=Comparator_Type))+
  geom_point(alpha=0.7)+
  facet_wrap(~year)+
  geom_smooth(method="lm", se = FALSE)
  
grad_debt_mdn_scatter
```

### Retention Over Time

```{r}

retention_regional <- ggplot(subset(jamovi_filtered, Comparator_Type %in% regional_comparators), aes(x=year_num, y=RET_FT4, color=INSTNM))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Retention Rate by Year for Regional Comparators"),x="Year (2000 to 2023)", y="Retention Rate (First-Time Students, in 4 Years)")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  xlim(2003,2023)
  
retention_regional

retention_national <- ggplot(subset(jamovi_filtered, Comparator_Type %in% national_comparators), aes(x=year_num, y=RET_FT4, color=INSTNM))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Retention Rate by Year for National Comparators"),x="Year (2000 to 2023)", y="Retention Rate (First-Time Students, in 4 Years)")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(,se=FALSE)+
  theme_light()+
  xlim(2003,2023)
  
retention_national

jamovi_filtered%>%
  filter(Comparator_Type %in% all_comparators)%>%
  ggplot(aes(x=year_num, y=RET_FT4, color=Comparator_Type, alpha=0.6))+
  geom_jitter(alpha=0.65)+
  labs(title=element_text("Retention Rate by Year"),x="Year (2000 to 2023)", y="Retention Rate (First-Time Students in 4 Years)")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_smooth(se=FALSE)+
  theme_light()+
  scale_color_manual(values=comparator_colors)+
  xlim(2003,2023)

```


```{r}
grad_debt_mdn_line <- ggplot(data=jamovi_filtered)+
  geom_line(aes(x=year_num,y=GRAD_DEBT_MDN))

grad_debt_mdn_line
```

### Linear Model for Graduate Debt + Residual Plot

```{r}
grad_debt_mdn_model <- lm(formula=COMP_ORIG_YR4_RT~GRAD_DEBT_MDN, data=jamovi_filtered)
summary(grad_debt_mdn_model)

s <- sqrt(sum(grad_debt_mdn_model$residuals^2)/174) #not sure what the 174 (originally 998)
s
```

```{r}
grad_debt_mdn_residual <- ggplot(jamovi_filtered, aes(x=GRAD_DEBT_MDN,y=grad_debt_mdn_model$residuals, alpha = 0.2))+
  geom_point()+
  geom_hline(yintercept=0, lty=2, lwd=1, color="blue")

grad_debt_mdn_residual # error?

qqnorm(grad_debt_mdn_model$residuals)
qqline(grad_debt_mdn_model$residuals)
```

### Histogram for Graduate Debt

- And completion rate for Pell...
- And retention rate for Pell...

```{r}
grad_debt_mdn_bar <- ggplot(data=jamovi_filtered)+
  geom_col(aes(x=as.numeric(COMP_ORIG_YR4_RT),y=as.numeric(GRAD_DEBT_MDN),fill="pos"))+
  facet_wrap(~year, scales="free_y")+
  labs(title="Title Here", subtitle="Subtitle Here", caption="Caption Here")+
  theme(legend.position="none",plot.title=element_text(face="bold",hjust=0.5),plot.caption=element_text(color="gray",size=5),plot.subtitle=element_text(hjust=0.5,size=8))
grad_debt_mdn_bar
```

